### 2017-03-28 ufpr/4-socialvanets/chapter4.tex

\chapter{Special properties of VANETs}
\label{chap:socialvanets}

%=====================================================

\section{Implications for trust management}

Despite the technological similarities between VANETs and MANETs, the differences in node behavior in VANETs creates the necessity of distinguished trust management solutions for those types of network.
Trust or reputation management systems in MANETs are usually developed around properties of MANETs such as: devices with constrained battery power, low-speed or no node mobility, and long contact times during each encounter.
As is described above, such properties are not valid for VANETs.

A few examples of trust management models for MANETs are shown below, and the reasons why they could not be applied to VANETs without significant modifications are explained.