### 2018-02-05 article/article.tex

\documentclass[conference]{IEEEtran}

\usepackage[utf8]{inputenc}	% arquivos LaTeX em Unicode (UTF8)
\RequirePackage[plainpages,pdfpagelabels]{hyperref}	% PDF com links, metadados

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  \graphicspath{{../pdf/}{../jpeg/}{./plots/}}
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
\fi

\usepackage{cleveref}

\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 
\def\scalecheck{\resizebox{\widthof{\checkmark}*\ratio{\widthof{x}}{\widthof{\normalsize x}}}{!}{\checkmark}}

% *** SPECIALIZED LIST PACKAGES ***
%
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\floatname{algorithm}{Algorithm}
\renewcommand{\algorithmiccomment}[1]{~~~// #1}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{TruMan: Trust Management for Vehicular Networks}

\author{\IEEEauthorblockN{Renan Greca}
\IEEEauthorblockA{Departamento de Informática\\
Universidade Federal do Paraná\\
Curitiba, Brazil\\
Email: rdmgreca@inf.ufpr.br}
\and
\IEEEauthorblockN{Luiz Carlos P. Albini}
\IEEEauthorblockA{Departamento de Informática\\
Universidade Federal do Paraná\\
Curitiba, Brazil\\
Email: albini@inf.ufpr.br}}

\maketitle

\begin{abstract}
By integrating processors and wireless communication units into vehicles, it is possible to create a vehicular ad-hoc network (VANET), in which cars share data amongst themselves in order to cooperate and make roads safer and more efficient.
A decentralized ad-hoc solution, which does not rely on previously existing infrastructure, Internet connection or server availability, is preferred so the message delivery latency is as short as possible in the case of life-critical situations.
However, as is the case with most new technologies, VANETs will be a prime target for attacks performed by malicious users, who may benefit from affecting traffic conditions.
In order to avoid such attacks, one important feature for vehicular networks is trust management, which allows nodes to filter incoming messages according to previously established trust values assigned to other nodes.
To generate these trust values, nodes use information acquired from past interactions; nodes which frequently share false or irrelevant data will have lower trust values than the ones which appear to be reliable.
This work proposes a trust management model in the context of daily commutes, utilizing the Working Day Movement Model as a basis for node mobility.
The results prove to be accurate and efficient, thanks to the low complexity of the algorithms constituting the trust model.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{section:introduction}

Within the next few years, a substantial share of new vehicles will come equipped with networking features \cite{connectedcar2016}.
These features will allow vehicles to quickly share data with other nearby devices and can be useful tools to reduce traffic and the risk of accidents.
Over 1 million people lose their lives to traffic accidents every year \cite{whotraffic}, so solutions to improve road safety are crucial for modern life.
%Vehicular ad-hoc networks are a much-studied usage of vehicular networking features.
%In these networks, all nodes are related to traffic; they can be vehicles equipped with on-board computers, or stationary units placed near roads.
By quickly sharing data with neighboring vehicles without the need of an Internet connection, smart vehicles can alert their drivers of important road conditions \cite{barba2012smart}, while autonomous vehicles can synchronize their movements to maximize traffic throughput \cite{amoozadeh2015platoon}.

The communication stardard for vehicular communication is the IEEE 802.11p or Wireless Access in Vehicular Environments (WAVE) \cite{jiang2008ieee}.
It describes two types of nodes for vehicular networks: on-board units (OBUs) and road-side units (RSUs).
Communication between two OBUs is called vehicle-to-vehicle (V2V) communication, while communication between an OBU and an RSU is called vehicle-to-infrastructure (V2I) communication.
This study focuses only on V2V scenarios, and therefore, any references to Vehicular ad-hoc networks (VANETs) and their nodes refer exclusively to vehicles with on-board units.

As is expected for new technologies, vehicular communications can become an appealing target for malicious users and attackers.
Some issues that could be exploited in such network include: vehicles with faulty sensors \cite{isaac2010security}; vehicles broadcasting false data \cite{golle2004detecting}; a flood of false data to generate a distributed denial of service (DDoS) scenario or to divert traffic \cite{garip2015congestion}; eavesdropping on other vehicles' communications, signal jamming or stalking \cite{isaac2010security}.

Each of these problems require specific solutions, although there are ways of making the network safer in general.
One way is taking advantage of the concept of trust between network members.
By having nodes remember previous interactions with one another, it is possible for them to build trust relationships and avoid those attacks that involve the spread of false data.
Trust solutions for VANETs are generally classified into data-oriented trust, which emphasizes the contents of a message, and entity-based trust, which emphasizes the sender of a message.
The solution described in this article is one of entity-based trust.

Using the proposed trust model, nodes in a vehicular network are able to identify which other nodes might be malicious.
Since the network is dynamic, nodes acquire more knowledge as time passes and the algorithm's results become more precise, taking advantage of social properties of VANETs \cite{da2013effective} \cite{cunha2014possible} to build strong relationships between nodes that connect frequently.
After running a series of simulations, the results were very promising; after one simulated days, nodes were able to correctly identify over 95\% of malicious nodes in the network.

%==> O QUE SE PRETENDE ATINGIR? O QUE TEM DE DIFERENTE? QUAIS RESULTADOS FORAM OBTIDOS? FALTA UM PARAGRAFO AQUI FALANDO DISSO.

The remainder of this paper is organized as follows.
Section \ref{section:previouswork} shows some examples of existing trust models for vehicular networks and the trust model for static networks that served as basis for this one;
\autoref{section:algorithm} explains the algorithms used to develop the trust model;
\autoref{section:results} shows the simulations which validate the usefulness of the trust model;
and \autoref{section:conclusion} presents closing thoughts on the project.

\section{Understanding the algorithm}
\label{section:algorithm}

%==> MUDAR O TITULO, PRACISA DIZER ALGO DO TIPO UNDERNEATH ALGORITHM

The objective of the Truman trust model is to allow nodes to infer whether or not other nodes in the network are malicious.
The algorithm that dictates the trust model runs continuously, with iterations happening in a preset interval.
In every iteration, a node collects information from its neighbors and runs a combination of algorithms to detect malicious nodes in the known network.
This information is maintained in a directed graph $T=(V,O)$, in which $V$ is the set of known vertices and $O$ is the trust relationship between pairs $u,v \in V$.
Graph $T$ is known as the \textit{trust graph}.

The edges $(u, v) \in O$ store a value between 0 and 1 that represents the degree of trust $u$ has for $v$.
This value is called the \textit{opinion} of $u$ about $v$.
At the start of the execution, edges have value $0.5$; this value increases when nodes have positive interactions and decreases otherwise.
A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.
By default, the threshold is $0.5$.

After collecting information from other nodes, Truman performs two steps: ($i$) it divides the network graph into strongly connected components using Tarjan's strongly connected components algorithm; ($ii$) it uses a graph coloring algorithm as a heuristic to determine which nodes to trust or not.

%In order to do this, a node must maintain a graph $T=(V,O)$ that represents its knowledge of the network, in which $V$ is the set of known nodes and $O$ is the set of opinions nodes have about others.
%Opinions are directed edges, each one with a weight between 0 and 1 that represents the degree of trust from one node towards another.
%A threshold $0 < h < 1$ may be used as a parameter to define the minimum weight for an edge to be considered positive, meaning that the origin node trusts the destination node.

The detailed descriptions of both algorithms are below, followed by the complete process of each iteration of the Truman trust model.

%==> VOCE AINDA NAO FALOU DOS PESOS DAS ARESTAS E COMO ELES FORAM CONSEGUIDOS
%==> VOCE NAO FALOU DOS GRAFOS

% ---------------------------------------------------------------------------------------------------
\subsection{Tarjan's strongly connected components algorithm}
\label{section:tarjan}
The use of Tarjan's strongly connected components algorithm \cite{tarjan1972depth} is an important aspect of Truman's efficiency.
This allows a large graph to be abstracted into a smaller one, which therefore reduces the input for further algorithms.
Given a directed graph $T = (V,O)$, a strongly connected component is defined as a group of nodes in which, for any pair of nodes $u, v \in V$, there exists a path from $u$ to $v$ and a path from $v$ to $u$.
For the purposes of trust management, this definition is extended to accept only paths of edges with weight above a predetermined threshold $h$.
Every node of the input graph $T$ must belong to a component.

In the implementation used, $index$, $lowlink$, $count$ and $stack$ are global variables accessible from every call of the function.
$index$ and $lowlink$ are arrays indexed by node IDs (predefined unique identifiers), $count$ is an integer and $stack$ is a last-in-first-out data structure.

The algorithm works by performing a depth-first search, adding nodes to $stack$ as they are visited.
If two nodes are present on $stack$, then there is a path from the first node to the second one (in the order they were added to $stack$).
Each node has two attributes assigned to it during the execution of the algorithm: $index$ is used to number the nodes in the order they are visited, while $lowlink$ is the lowest indexed node reachable from each node.

In the call that visits a node $u$, the algorithm must loop through each node $v$ trusted by $u$ (that is, $u \rightarrow v$ exists and has value greater than $h$).
If node $v$ has not yet been visited, the algorithm is called for  $v$.
The $lowlink$ of $u$ is then calculated as the smallest value between $lowlink[u]$ and $lowlink[v]$, because any node reachable from $v$ is also reachable from $u$.
After the loop, if $lowlink[u]$ is equal to $index[u]$, it means that $u$ is the lowest indexed node reachable from itself and that it is the root of a component.
Therefore, nodes must be popped from the $stack$ until $u$ is found.
Each node popped, including $u$, is a member of a strongly connected component.

The number of components is, at most, $|v|$.
In a worst-case scenario, each node is put into its own component; however, this would not be the case for most useful input graphs.

Algorithm \autoref{algorithm:tarjan} shows the general structure of Tarjan's algorithm  \cite{tarjan1972depth}.
The complexity of the algorithm is $O(|V|+|O|)$ for a graph $T = (V,O)$.

With the results of Tarjan's algorithm, an undirected component graph $C = (V',O')$ is formed.
Each $v' \in V'$ is the abstraction of one component identified by Tarjan's algorithm, while the edges $o' \in O'$ are edges from $T$ between nodes that do not belong in the same component.

\begin{algorithm}
\caption{Tarjan's strongly connected components algorithm}\label{algorithm:tarjan}
\begin{algorithmic}[1]

\Function{Tarjan}{vertex $u$}

\State $index[u] = count$
\State $lowlink[u] = count$
\State $count \gets count + 1$
\State \textbf{push} $u$ \textbf{to} $stack$

\For{$v$ \textbf{in} neighbors of $u$ }
	\If{weight of $u \rightarrow v$ $<$ $h$}
		\State \textbf{continue}
	\EndIf

	\If{$index[v] = -1$}
		\Comment{\emph{v has not been visited yet}}
		\State \textbf{Tarjan}($v$)
	\EndIf

	\State $lowlink[u] \gets \textbf{min}(lowlink[u],lowlink[v])$
\EndFor

\If{$lowlink[u] = index[u]$}
	\Repeat
		\Comment{\emph{unstack nodes until u is found}}
		\State \textbf{pop} $w$ \textbf{from} $stack$
		\State \textbf{add} $w$ \textbf{to} $component$
	\Until {$w = u$}
\EndIf

\EndFunction
\end{algorithmic}
\end{algorithm}

% ---------------------------------------------------------------------------------------------------
\subsection{Graph coloring with minimum colors}
\label{section:coloring}
The algorithm proposed in \cite{mittal2011graph} is an efficient approach to graph coloring.
Graph coloring is one of the possible heuristics used to detect malicious nodes after the generation of the component graph using Tarjan's algorithm.
Out of the tested heuristics, it presented the best results, so it has been chosen as the heuristic for the trust model.

%The process of graph coloring consists of labeling each node with a color so that no two neighboring nodes share the same color.
%This problem has been studied in Computer Science since, at least, 1972 \cite{karp1972reducibility} and has been studied as a classic mathematics problem for even longer \cite{kempe1879geographical}.
It has been proven mathematically that any planar graph can be colored with at most four colors \cite{appel1976every}, but discovering the smallest number of colors necessary to color an arbitrary graph (called the graph's chromatic number) is an NP-hard problem \cite{sanchez1989determining}.
In \cite{mittal2011graph}, the authors propose to color a graph using the minimum possible amount of colors.
Although they do not prove that their algorithm always uses the smallest possible amount of colors, the output is always a correct coloration and the algorithm is nevertheless efficient.
The complexity of the algorithm is $O(|E|)$ for an undirected graph $G = (V,E)$.
As a comparison, the classic DSATUR algorithm for graph coloring has complexity $O(|V|^2)$ \cite{brelaz1979new}.
%For the purposes of this study, it is not necessary to prove that the coloring algorithm's output uses the minimum possible number of colors.

Algorithm \autoref{algorithm:coloring} shows the general structure of the graph coloring algorithm  \cite{mittal2011graph} \cite{vernize2013dissertation}.

A limitation of this algorithm is that the edges must be sorted according to node indexes beforehand.
It does not matter which nodes get assigned which indexes, but once they are assigned those numbers, the algorithm must follow the edges in numerical order.
This is demonstrated in \cite{vernize2013dissertation}.

\begin{algorithm}
\caption{Graph coloring with minimum colors}\label{algorithm:coloring}
\begin{algorithmic}[1]

\Function{Coloring}{graph $G$}

\State \textbf{color} all nodes of $G$ \textbf{with} 0
\State $d \rightarrow 0$
\For{$e=(u,v)$ \textbf{in} edges of $G$}
	\If{$u$ and $v$ have the same color}
		\If{$color[v] = d$}
			\State $d \rightarrow d+1$
		\EndIf
		\State $color[v] \rightarrow d$
	\EndIf
\EndFor

\EndFunction
\end{algorithmic}
\end{algorithm}

% ---------------------------------------------------------------------------------------------------
\subsection{The Truman algorithm}
\label{section:trustmanagement}
%\subsection{Malicious Node Identification Algorithm}
%\label{section:mani}

%The basis of the presented trust model is the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, \cite{vernize2013dissertation}.
%The authors present a malicious node identification scheme based on strongly connected components and graph coloring.
%The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
%Furthermore, the algorithm is executed by a global observer which has information about the complete network.

Truman is based on the MaNI algorithm \cite{vernize2015malicious}, which suggested the use of Tarjan's algorithm and the graph coloring algorithm used here.
However, MaNI was developed for static networks such as social networks, and is executed by a supervising agent, making it unsuitable for a vehicular network.
The work describes here takes advantage of the efficiency of the two graph theory algorithms into a model that is viable for highly dynamic networks.

In order to work with dynamic networks, the Truman algorithm must consider snapshots as inputs.
The algorithm runs continuously, running new iterations at a predetermined interval.
Each iteration $i$ is associated with a timestamp, which indicates when the snapshot was taken.
A snapshot $G_i = (V_i, E_i)$ represents the complete topology of the network at the iteration $i$; $V_i$ is the set of nodes participating in the network at that time and $E_i$ is the set of edges connecting pairs of nodes within communication range of each other.
As the network is dynamic, it is expected that each $G_i$ is different from $G_{i-1}$, but also that the changes follow a determined mobility model.

Furthermore, the algorithm runs in a decentralized fashion, meaning each node in the network runs its own instance of the algorithm.
%This is necessary because it is not reasonable to expect a supervisor of the network to have complete knowledge of the nodes and relationships in the network.
Each node starts knowing information only about itself and maintains its own abstraction of the network surrounding it.
Every node $u$ has a static, connected and directed trust graph $T^u = (V^u, O^u)$, in which $V^u$ is the set of nodes $u$ is aware of and $O^u$ is the set of trust relationships (opinions) $u$ knows between members of $V^u$.
Since $T^u$ changes over time, there is a $T^u_i = (V^u_i, O^u_i)$ for every iteration $i$.

%==> VOCE NAO FALOU DOS PESOS DAS ARESTAS AINDA

%==> ACHO QUE PRECISA FALAR AQUI DA PRINCIPAL DIFERENCA PARA O MaNI, QUE AQUI ESTAMOS FALANDO DE REDE MOVEL E ALGORITMO DISTRIBUIDO

At first, the node is collecting and organizing information.
The storage of an abstraction of the network in the form of a trust graph $T$ is crucial for the remainder of the algorithm.
A prerequisite of step 2 is a test that correctly classifies a neighboring node as benign or malicious.
After these steps, $T^u_i$ is formed, which is then used for the following steps.

Each edge $w \rightarrow v \in O^u_i$ is weighted according to the degree of trust $w$ has for $v$ with a value between 0 and 1.
The closer the value is to 1, the more $w$ trusts $v$.
These values are not mutual, so the value of $w \rightarrow v$ can be different from the value of $v \rightarrow w$.

After the collection of data, $T^u_i$ is separated into strongly connected components using Tarjan's algorithm \cite{tarjan1972depth}, which is described in detail on \autoref{section:tarjan}.
For each node in a component, there is a path formed by edges of weight higher than a threshold $0 < t < 1$ to each other node in the same component.
In other words, within a single component, all nodes trust one another directly or indirectly; nodes that do not satisfy this condition are separated into different components.
Each of these components becomes a node of a component graph $C^u_i = (V'^u_i, O'^u_i)$.

The creation of $C^u_i$ simplifies the remaining computation.
Since each vertex $v' \in V'^u_i$ is a component of $T^u_i$ in which all nodes trust each other, for the purposes of identifying malicious nodes, all nodes within each of those components can be treated as the same.
They can either be benign nodes which legitimately trust one another, or malicious nodes colluding with each other.
After the formation of $C^u_i$, one or more heuristics can be used to classify the nodes as benign or malicious.

The authors of \cite{vernize2015malicious} describe the coloring heuristic, which can use a graph coloring algorithm such as DSATUR \cite{brelaz1979new} or the algorithm described in \autoref{section:coloring} \cite{mittal2011graph}.
After running either algorithms with graph $C^u_i$ as input, the color whose nodes in $C^u_i$ represent the most nodes in $T^u_i$ is classified as correct, and all others are classified as malicious.
Once this information in $C^u_i$ is brought back to graph $T^u_i$, it is trivial to label the nodes in $T^u_i$ as either benign or malicious based on their components' classifications.

In the experiments shown in \cite{vernize2013dissertation}, the coloring heuristic shows the most promising results, identifying a high ratio of the malicious nodes in the network.
Other heuristics were experimented with, but were either less effective in detecting malicious nodes, or provided too many false positives.
Therefore, for the purposes of this research, only the coloring heuristic is considered.

In summary, every node $u$ runs the following steps in each iteration to detect malicious nodes in the network:

\begin{enumerate}
	\item Node $u$ checks who are its neighbors (nodes within its communication range).
		  Newly discovered nodes and newly formed edges are added to $T^u_i$.
		  Edges are created with weight 0.5.
	\item Node $u$ tests all of its neighbors to discover which ones can be directly trusted or not.
		  New trust values are computed for the edges using the average between the previous value and either 1 (if the neighbor is trustworthy) or 0 (otherwise).
	\item If a neighbor $v$ is trustworthy, $u$ merges $T^v_{i-1}$ into $T^u_i$.
	\item Tarjan's algorithm is executed to identify the strongly connected components of $T^u_i$, resulting in a component graph $C^u_i$.
	\item The graph coloring algorithm is executed on $C^u_i$ and nodes are classified as benign or malicious. %according to the same rules as the MaNI algorithm.
\end{enumerate}

%Two types of experiments were made in each network: first, all malicious nodes inverted the edge weights leading to their neighbors; second, malicious nodes randomly inverted or not the weights.
%In the first scenario, the results show excellent precision in most networks, detecting nearly every malicious node.
%Experimenting with the second scenario, the results are less precise, however still promising: with up to 20\% of malicious nodes in the network, the error rate is under 7\%, while with the worst case, 50\% of the network being malicious, the error rate is approximately 15\%.
%
%The authors suggest running the algorithm repeatedly after removing the malicious nodes from the network.
%By doing this, nearly all malicious nodes are detected by it even when randomly changing edge weights.

% ---------------------------------------------------------------------------------------------------

%The MaNI algorithm works well and is efficient for static graphs.
%However, some important modifications had to be made to accommodate dynamic graphs in a decentralized fashion.
%The main changes made are described below.

%First of all,

%Another important change is on the trust values themselves.
%MaNI uses binary trust (either 0 or 1).
%This has been replaced by a floating-point number between 0 and 1; a predefined threshold defines which values are deemed trustworthy or not.
%This was not the case in the first versions of the trust model; it was added in order to avoid abrupt changes to opinions.
%Since the network is different on each iteration of the algorithm, there were instances in which a change to the topology caused a radical change to the output of the graph coloring algorithm.
%Using a trust value that gradually changes over time, output variations between two iterations became small and gradual.
%
%At the start of an execution, a node knows only about itself.
%From that point, iterations of the algorithm will run in a set interval of $l$ seconds.
%In each iteration (labeled $t$ after the timestamp it occurs on) and for every node $u$, the following steps are executed:

%To mention:
%- Graph building;
%- Change from binary trust to a weighted model;
%- Evolution of trust over time.

%First of all, there is an important change to the graph knowledge used to run the algorithm.
%The complete trust graph is a parameter to the MaNI algorithm, which is executed by a supervisor with global knowledge of the graph.
%
%In this case, given a complete trust graph $T = (V, O)$, each node $u \in V$ has its own vision of the network and has to build a graph, called $T_u$, to represent its surrounding network.
%In each iteration of the algorithm, $u$ receives information from its neighbors about their own graphs.
%For every neighbor $v \in V$, $T_u$ must me merged with $T_v$.
%An edge $u \rightarrow v$ exists if $u$ is aware of $v$.
%Given sufficient iterations, $T_u$ forms a complete or nearly complete vision of the network.

%Alongside $G$, there is the trust graph $T = (V, O)$ which stores trust relationships between pairs of nodes in the form of directed edges.
%Again, each node $u$ has its own $T_u$, which contains the trust information node $u$ knows about.
%When a node $u$ asks for network information from node $v$, it also merges $T_u$ with $T_v$, but only if $u$ trusts $v$.
%If $u$ deems $v$ malicious, any information originating from $v$ cannot be considered valid.

%Furthermore, this implies that the malicious node detection algorithm runs using incomplete graph knowledge.
%This is expected and generally unavoidable in vehicular networks, which have high mobility and can be extremely large.

\section{Results}
\label{section:results}

In order to test the trust model, simulations were made using an implementation of the algorithm in Python.
To generate the input graphs with node mobility, the ONE simulator \cite{keranen2009one} was used in conjunction with the Working Day Movement Model \cite{ekman2008working}, which provides a reasonable imitation of vehicle movement in real life.
Snapshots of the network were taken every 10 simulated seconds, and these snapshots were used as input for the algorithm.

Most of the parameters for the simulator were taken from the article detailing the Working Day Movement Model.
The simulation ran for 86400 seconds (24 hours), with a work day length of 28800 seconds and a standard deviation of departure time of 7200 seconds.
Nodes move between 7 and 10 m/s in an area of approximately 14 km\textsuperscript{2} based on a section of the map of Helsinki.
%The number of nodes and their communication range vary for different simulations.
There is a total of 160 nodes, 150 of which are following the Working Day Movement Model, and the other 10 are moving randomly to simulate vehicles that do not follow daily patterns.
Since this simulation is for vehicles instead of pedestrians, there are no buses in the model and every node is guaranteed to own a vehicle and travel by car.
The parameters regarding offices, meeting spots and shopping were kept intact.

\subsection{Network Density}
The communication range of nodes vary from 10m to 50m, to illustrate the impact of different network densities.
The network density is a value that abstracts the volume and frequency of connections in a vehicular network.
It is calculated using the communication range of the nodes, the amount of nodes, and the total area of the simulation.
For this trust model, higher densities yield better results, since nodes can construct and update their models of the network more quickly (this is demonstrated in \autoref{subsection:simulations}).

The simulations shown here have densities that vary between 0.001 (10m range) and 0.04 (50m range).
As a comparison, the density of São Paulo was calculated as 2.24 with 10m range, a value much higher than what is necessary for the algorithm.

\subsection{Simulations}
\label{subsection:simulations}
To improve readability, all figures in this section follow the same format.
The $X$ axis shows the results of sequential iterations, ranging from 0 to 8639, while the $Y$ axis shows a percentage of all nodes in the network, ranging from 0 to 100.
The blue line represents the percentage of nodes detected out of the complete network.
Magenta is the percentage of malicious nodes in the network (ground truth).
Finally, green represents the nodes correctly identified as malicious (true positives), cyan represents the undetected malicious nodes (false negatives) and red represents the benign nodes incorrectly identified as malicious (false positives).

\autoref{fig:random103050} shows the results of simulations running with 10\% of nodes acting maliciously, with communications range varying from 10m to 50m.
It is possible to see how the increase in communication range allows the algorithm to converge much sooner, taking over 8000 iterations with 10m range and achieving solid results at just over 1000 iterations with 50m range.

\autoref{fig:random0} and \autoref{fig:random1} show the variation of results for different amounts of malicious nodes in the network.
By the end of one day, the algorithm is able to detect all malicious nodes when they are up to 30\% of the network.
At 40\%, a small part of malicious nodes are yet to be detected.
At 50\%, as is expected, the results are inconsistent as control of the network is completely divided between benign and malicious nodes; at this point, the network is completely compromised.
The amount of malicious nodes also affects network discovery, since nodes do not trust information from malicious neighbors.

\autoref{fig:random7} shows the execution of the algorithm over the course of 7 days.
Most malicious nodes are identified by the end of the first day; in the following iterations, the algorithm finishes building the network model and sorts out remaining false negative or false positive results.
After iteration 20000, the results are completely consistent.

%==> ACHO QUE PODIA USAR O MULTIFIGURE PARA AGRUPAR UM POUCO AS FIGURAS. COLOCA 3 LADO A LADO. ASSIM SOBRE ESPAÇO NO FINAL PARA PODERMOS COMPARAR COM TRABALHOS RELACIONADOS.

% ----
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\lin0.5ewidth]{Network_A_10.0/new_plots/10.png}
%%\includegraphics[width=0.45\linewidth]{Network_A_10.0/new_plots/30.png}
%\caption{Simulation without random nodes and range of 10m.} \label{fig:nonrandom1}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_A_10.0/new_plots/30.png}
%\caption{Simulation without random nodes and range of 30m.} \label{fig:nonrandom2}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_A_10.0/new_plots/50.png}
%\caption{Simulation without random nodes and range of 50m.} \label{fig:nonrandom3}
%\end{figure}

% ----

\begin{figure*}
\centering
\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/30.png}
%\caption{Simulation with random nodes and range of 10m.} \label{fig:random10}
\hspace{1pt}
\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/30.png}
%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
\hspace{1pt}
\includegraphics[width=0.3\linewidth]{Network_rA_10.0/new_plots/50.png}
\caption{Simulation with random nodes, 10\% malicious and range of 10m, 30m and 50m.} 
\label{fig:random103050}
\end{figure*}

%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/30.png}
%\caption{Simulation with random nodes and range of 30m.} \label{fig:random30}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_rA_10.0/new_plots/50.png}
%\caption{Simulation with random nodes and range of 50m.} \label{fig:random50}
%\end{figure}

% ----

\begin{figure*}
\centering
\includegraphics[width=0.3\linewidth]{Network_rA/10_1}
\hspace{1pt}
\includegraphics[width=0.3\linewidth]{Network_rA/10_5}
\hspace{1pt}
\includegraphics[width=0.3\linewidth]{Network_rA/10_10}
\caption{Simulation with random nodes, range of 10m and 1\%, 5\% or 10\% malicious.}
\label{fig:random0}
\end{figure*}

%\begin{figure}
%\centering0.5
%\includegraphics[width=0.5\linewidth]{Network_rA/10_5}
%\caption{Simulation with random nodes, range of 10m and 5\% malicious.} \label{fig:random1}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_rA/10_10}
%\caption{Simulation with random nodes, range of 10m and 10\% malicious.} \label{fig:random2}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_rA/10_20}
%\caption{Simulation with random nodes, range of 10m and 20\% malicious.} \label{fig:random3}
%\end{figure}

\begin{figure*}
\centering
\includegraphics[width=0.3\linewidth]{Network_rA/10_30}
\includegraphics[width=0.3\linewidth]{Network_rA/10_40}
\includegraphics[width=0.3\linewidth]{Network_rA/10_50}
\caption{Simulation with random nodes, range of 10m and 30\%, 40\% or 50\% malicious.}
\label{fig:random1}
\end{figure*}

%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_rA/10_40}
%\caption{Simulation with random nodes, range of 10m and 40\% malicious.} \label{fig:random5}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{Network_rA/10_50}
%\caption{Simulation with random nodes, range of 10m and 50\% malicious.} \label{fig:random6}
%\end{figure}

% ----

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{Network_rA7/10_10}
\caption{Simulation with random nodes, range of 10m and 10\% malicious during 7 days.} \label{fig:random7}
\end{figure}

%\begin{subfigure}{0.20\textwidth}
%\includegraphics[width=\linewidth]{images/coloring/4.png}
%\caption{Edge $(2,3)$ is checked and node $3$ gets the current value of $d$.} \label{fig:coloring5}
%\end{subfigure}
%\hspace*{2cm} % separation between the subfigures
%\begin{subfigure}{0.20\textwidth}
%\includegraphics[width=\linewidth]{images/coloring/5.png}
%\caption{Edge $(2,4)$ is checked and node $4$ gets the current value of $d$.} \label{fig:coloring6}
%\end{subfigure}
%\hspace*{2cm} % separation between the subfigures
%\begin{subfigure}{0.20\textwidth}
%\includegraphics[width=\linewidth]{images/coloring/6.png}
%\caption{Edge $(3,4)$ is checked and node $4$ gets a new color. $d=4$ (yellow).} \label{fig:coloring6}
%\end{subfigure}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Previous work}
\label{section:previouswork}

==> FALTA DETALHAR ELES.

Several models have been proposed to solve the problem of trust in vehicular networks. 
In this section, some of the most relevant ones are described, considering the time in which they were proposed, the advantages they bring and their contributions to later study. 
None of them provide a complete solution, but serve as pieces of a puzzle that is still incomplete. 
Many trust management solutions for VANETs have been proposed over the years, such as \cite{patwardhan2006data}, \cite{gerlach2007trust}, \cite{raya2008data}, \cite{huang2010situation}, \cite{ding2013novel}, \cite{haddadou2013trust}, \cite{liu2016lsot}, \cite{kerrache2016detection}.
There are also some review and/or survey articles on the subject of VANET trust models, such as \cite{zhang2011survey}, \cite{ma2011survey}, \cite{zhang2012trust}, \cite{mejri2014survey}, \cite{soleymani2015trust} \cite{sengar2016survey}, and \cite{dwivedi2016review}.

For the Malicious Node Identification Algorithm (MaNI) proposed in \cite{vernize2015malicious}, the authors present a malicious node identification scheme based on strongly connected components and graph coloring.
The model is proposed for complex networks in general, but is not suited for VANETs because it is designed only for static networks.
Furthermore, the algorithm is executed by a global observer which has information about the complete network.
It is, however, very efficient thanks to the classification of nodes into components and the usage of a fast heuristic.
The usage of components and coloring serve as basis for the trust model proposed here, which is expanded to work on distributed and dynamic networks such as vehicular networks.

\begin{table}[h!]
\caption{Properties of Truman and related work}
\label{table:schedule}
\centering
\begin{tabular}{|p{1.5cm}||p{1.5cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|p{0.4cm}|}
 \hline
 \textbf{Task} & Truman & \cite{vernize2013dissertation} & \cite{golle2004detecting} & \cite{patwardhan2006data} & \cite{minhas2010towards} \\
 \hline
 Decentralized & - & \checkmark & - & - & \\
 \hline
 Sparsity & - & \checkmark & - & - & \\
 \hline
 Dynamics & - & \checkmark & - & - & \\
 \hline
 Scalability & - & \checkmark & - & - & \\
 \hline
 Confidence & - & \checkmark & - & - & \\
 \hline
 Security & - & \checkmark & - & - & \\
 \hline
 Privacy & - & \checkmark & - & - & \\
 \hline
 Robustness & - & \checkmark & - & - & \\
 \hline
 Cost & $O(|V|* |O|)$ & n/a & n/a & n/a & n/a \\
 \hline
\end{tabular}
\end{table}

\section{Conclusion}
\label{section:conclusion}

While malicious nodes can diminish the usefulness of vehicular networks, the concept of trust as applied in VANETs is a powerful tool for those seeking to reduce the spread of false information as much as possible.
In this paper, a new trust model for vehicular networks was presented, which combines the efficiency of previous algorithms in order to generate fast and accurate results.

As nodes travel across the network and collect more data from neighbor, they are able to form an abstraction of the network which can be used to detect malicious nodes.
By placing nodes into strongly connected component with Tarjan's algorithm, a network containing a large amount of node can be simplified into a much smaller one.
And, using a simple graph coloring algorithm, most malicious nodes stand out by having colors different than the majority of nodes.

The experiments show that vehicles within a network can form a sufficient model of the network in around one day, and by then they are also able to detect nearly every malicious node in the network, with a small amount of false positives.
Since trust amongst nodes in the network is a value ranging between 0 and 1, the trust between two nodes gradually increases or decreases according to the results of each iteration.

% conference papers do not normally have an appendix

% use section* for acknowledgment
%\section*{Acknowledgment}
%
%
%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{article}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}