### 2018-04-17 dissertation/0-preambulo/resumoextendido.tex

\begin{resumoextendido}

	\section*{Introdução}
	Este é o resumo extendido em português.

	\section*{Redes Complexas}

	\section*{Redes Veiculares}

	\section*{TruMan}

	\section*{Simulações}

	\section*{Conclusão}
\end{resumoextendido}